---
title: "Triangulating iconicity: Iconicity ratings"
author: "Mark Dingemanse & Stella Punselie"
date: "Updated `r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---
  
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='figures_md/',
                      echo=TRUE, warning=FALSE, message=FALSE)
```

Code notebook for a study of the relation between iconicity ratings and experimentally collected guessability scores.

## Setup

```{r preliminaries, results="hide"}
# Packages
library(tidyverse)
library(readxl)
library(plyr)
library(ggthemes)
library(gghalves)
library(car)
library(emmeans)
library(plotly)
library(ggpubr)
library(viridis)

# useful functions
`%notin%` <- function(x,y) !(x %in% y) 
mean.na <- function(x) mean(x, na.rm = T)
sd.na <- function(x) sd(x, na.rm = T)

```

## Rating: descriptives

```{r}
# load dataframe with all ratings
read.csv("data\\ideophones_rated_uncorr.csv",header=TRUE,na.strings=c("", "NA"),sep=";",check.names=FALSE) -> d.uncorr

# to long format
d.uncorr <- gather(d.uncorr, item, rating, 6:245)
d.uncorr <- d.uncorr[,c(1,6,7)]
colnames(d.uncorr) <- c("rater","item","rating")
#remove empty cells
d.uncorr <- subset(d.uncorr, !is.na(d.uncorr[,"rating"]))
#change ratings to numeric values with decimals
as.numeric(sub(",", ".", d.uncorr$rating, fixed = TRUE)) -> d.uncorr$rating


# check for inconsistent raters with person-total correlation based on Motamedi et al. (2019)

ptc <- d.uncorr[,c("item","rater","rating")]
personTotalCorrelationCorrected <- function(ptc)
  {
  require(tidyverse)
  ptc %>%
    dplyr::rename(raterFocal = rater,
           ratingFocal = rating) %>%
    full_join(ptc, by='item') %>%
    filter(rater!=raterFocal) %>%
    group_by(item, raterFocal, ratingFocal) %>%
    summarise(ratingsOthers = mean(rating, na.rm=TRUE)) %>%
    base::split(.$raterFocal) %>%
    map(~cor(.$ratingFocal, .$ratingsOthers, 
             use="pairwise.complete.obs")) %>%
    as.data.frame %>%
    gather %>%
    dplyr::rename(rater = key, perTotCor = value)
}
View(personTotalCorrelationCorrected(ptc))

#plot person-total correlations
ptc %>%
select(item, rater, rating) %>%
personTotalCorrelationCorrected %>%
ggplot(aes(x=perTotCor)) + 
  geom_density() + 
  geom_histogram(stat="bin", 
                 alpha=0.4,
                 aes(y=..density..),
                 binwidth=0.1) +
  labs(x="Person-total correlation")
ggsave("figures\\rating_PerTotCorr.png",height=5,width=7.5)
```

One participant (N045) showed a negative person-total correlation, which according to Curran (2016) indicates a careless responder rendering invalid data. Inspection of this participant’s data showed they had indeed given the same answer to every question. I therefore decided to remove this participant from the dataset.

Two other participatns (N022 and N034) also had to be removed because of their knowledge of Japanese and Korean, resulting in a total of 75 participants to be included in the analysis.

```{r}
# load corrected data
d <- read_xlsx("data\\ideophones_rated.xlsx")
d.means <- read_xlsx("data\\ideophones_rated_means.xlsx")

# get means and standard deviations across all ideophones and for each language and category
mean(d$rating)
  # 2.948339
sd(d$rating)
  # 1.297106
sumcat <- ddply(d,~category,summarise,mean=mean(rating),sd=sd(rating))
sumlang <- ddply(d,~language,summarise,mean=mean(rating),sd=sd(rating))

```

Because the category "Other" is quite a vague residual category with only 17 items (as opposed to 30-40 items for the other categories), I decided to remove this category for now. I will keep a copy that includes the ratings for "Other".

```{r}
#create copy of data including the category "Other"
d.other <- read_xlsx("data\\ideophones_rated.xlsx")
d.m.other <- read_xlsx("data\\ideophones_rated_means.xlsx")

#remove category "Other" from d and d.means
d <- d[!grepl("Other",d$category),]
d.means <- d.means[!grepl("Other",d.means$category),]

# make new summary for language (without the Japanese "other" ideophones)
sumlang2 <- ddply(d,~language,summarise,mean=mean(rating),sd=sd(rating))
```

Here we plot the distribution of the rating data.

```{r}
#convert some columns to factors
cols <- c("category","language","study")
d[cols] <- lapply(d[cols], factor)
d.means[cols] <- lapply(d.means[cols], factor)

#combined boxplot and scatterplot by category
ggplot(d.means, aes(x=reorder(category,-rating),y=rating,color=category)) +
  theme_tufte(base_size=16) +
  geom_half_boxplot(aes(middle=mean(rating)),show.legend=F) +
  geom_half_point(show.legend=F) +
  scale_colour_viridis_d(option="D",alpha=0.8) +
  xlab("category") + scale_x_discrete(labels = c("Sound","Motion","Shape", 
  "Texture","Colour/Visual"))
ggsave("figures\\rating_category.png",height=5,width=7.5)

#same, but by language
ggplot(d.means, aes(x=reorder(language,-rating),y=rating,color=language)) +
  theme_tufte(base_size=16) +
  geom_half_boxplot(aes(middle=mean(rating)),show.legend=F) +
  geom_half_point(show.legend=F) +
  scale_colour_viridis_d(option="D",alpha=0.8) +
  xlab("category")
ggsave("figures\\rating_language.png",height=5,width=7.5)

```

## Statistics

Now we want to see which categories and languages differ significantly from the others in their iconicity ratings.

First, we check the assumptions.

```{r}
# check normality by visual inspection
ggqqplot(d, "rating", facet.by = "category")
  # all points fall approximately along the reference line, so normality can be assumed

# check homogeneity of variance with Levene's test
leveneTest(rating~category,data=d)
leveneTest(rating~language,data=d)
  # both are not significant (> .05), so homogeneity of variance can be assumed
```

Now we construct a linear model with category as a predictor of the iconicity ratings.

```{r}
summary(lm.cat <- lm(rating ~ category, d))

# pairwise comparisons with Tukey adjustments in the emmeans package
emm.cat <- emmeans(lm.cat, "category")
pairs(emm.cat)
pwpm(emm.cat)
```

And here we do the same with language as predictor.

```{r}
summary(lm.lang <- lm(rating ~ language, d))

# pairwise comparisons in emmeans package
emm.lang <- emmeans(lm.lang, "language")
pairs(emm.lang)
pwpm(emm.lang)
```

## Combined data
Here we load the combined coding and guessability data and then add the rating data.

```{r}
# get coded and guessed data from previous step (TI_02_congruence)
d = read.csv("data\\ideophones_coded_guessed.csv")

coding_categories <- d %>% dplyr::select(matches('F_|M_'),-matches('notes|meaning')) %>% names()

# add mean ratings
d.ratings = read_excel("data\\ideophones_rated_means.xlsx") %>%
  dplyr::select(-c(item,category,list))
d <- left_join(d,d.ratings,by=c("filename","language","study"))

# convert ratings to z-scores for comparison with guessability (logodd) scores
d$rating_z <- scale(d$rating,center=T,scale=T)

#remove category "Other"
d <- d[!grepl("Other",d$category),]

```

## Visualisations
Let's first visualise the relation between the rating and guessing data.

```{r}
ggplot(d,aes(x=rating_z,y=score_z)) +   
  geom_point(aes(colour=study),position="jitter") +
  geom_smooth(method=lm,colour="black") +
  theme_tufte(base_size = 16) +
  xlab("rating") + ylab("guessability")
#ggsave("figures\\rating_guessing_corr.png",height=5,width=7.5)

ggplot(d,aes(x=rating_z,y=score_z)) +   
  geom_point(aes(colour=category,shape=study),position="jitter") +
  geom_smooth(method=loess,colour="black") +
  theme_tufte(base_size = 16) +
  xlab("rating") + ylab("guessability")
```

There appears to be a positive correlation between the rating and guessing data.

Now let's visualise the relations between the coding, guessing and rating data.

```{r problem chunk}

ggplot(d,aes(x=rating_z,y=logodds, color=C_cumulative)) +
  geom_point() +
  geom_smooth(method=loess) +
  scale_fill_viridis(option="plasma",discrete=F,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=F,begin=0.3,end=0.9) +
  NULL

```


## Statistics
```{r}
# Pearson correlation
cor.test(d$score_z,d$rating_z,method="pearson")
cor.rank <- cor.test(d$score_z,d$C_cumulative,method="spearman")
```

Pearson's correlation coefficient indicates that there is indeed a strong positive correlation between the guessing and rating data, r(220) = .61, p < .001.

```{r}
#check where the guessing and rating data deviate
d %>%
  filter(rating_z-score_z > 2 | rating_z-score_z < -2)%>%
  dplyr::select(filename,ideophone,rating,rating_z,language,logodds,score,score_z) %>%
  arrange(score) %>% ungroup()
```

